---
title: "gtsummary in quarto"
format:
  html: #change this to pdf if want to render as such need "quarto install tinytex" in terminal first
    theme: quartz
execute: 
  eval: true
  echo: false
  error: true #if you put true this will still render the doc with the error message 
  
---

We’re using some data from the National Longitudinal Survey of Youth 1979, a cohort of American young adults aged 14-22 at enrollment in 1979. They continue to be followed to this day, and there is a wealth of publicly available data online. I’ve downloaded the answers to a survey question about whether respondents wear glasses, a scale about their eyesight with glasses, their (NLSY-assigned 😒) race/ethnicity, their sex (presumably at birth, but unclear), their family’s income in 1979, and their age at the birth of their first child.

```{r}
#| output: false #makes the code show but not the results 
# #| message: false #no messages from packages but will show nrows function output

library(tidyverse)
library(gtsummary)
nlsy_cols <- c("glasses", "eyesight", "sleep_wkdy", "sleep_wknd", 
               "id", "nsibs", "samp", "race_eth", "sex", "region", 
               "income", "res_1980", "res_2002", "age_bir")
nlsy <- read_csv(here::here("data", "raw", "nlsy.csv"), 
                 na = c("-1", "-2", "-3", "-4", "-5", "-998"),
                 skip = 1, col_names = nlsy_cols) |> 
  mutate(region_cat = factor(region, labels = c("Northeast", "North Central", "South", "West")),
         sex_cat = factor(sex, labels = c("Male", "Female")),
         race_eth_cat = factor(race_eth, labels = c("Hispanic", "Black", "Non-Black, Non-Hispanic")),
         eyesight_cat = factor(eyesight, labels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
         glasses_cat = factor(glasses, labels = c("No", "Yes"))) 

```

This is the number of rows in the table:
```{r}
#| label: Outline label for this chunk
nrow(nlsy)
```

# Customization of `tbl_summary()`

```{r}
#Choose a table to label and caption, and then write a sentence that cross-references it (e.g., Table 1 shows the descriptive statistics)

#| label: tbl-one
#| tbl-cap: "First Table"
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

@tbl-one shows the descriptive statistics.


```{r}
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    region_cat ~ "Region",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing")
```

## Exercise 2.2 (Part 2)

```{r}
tbl_summary(
  nlsy, #NLSY - object not being found cause of the CAPS name,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

## Exercise 2.3 (Part 1)
```{r}
#Add a chunk where you calculate the mean number of hours of sleep on weekends (you can use the mean() function). Have the code print that output.
# Now, store that value as an object in the chunk. Add it as inline text.
# Finally, calculate the statistic directly as inline text.


#| output: true
avg_sleepwknd <- mean(nlsy$sleep_wknd, na.rm = TRUE)
```

-   Participants' mean number of hours of sleep on weekends is `r avg_sleepwknd`. 
-   I am calculating the mean number of hours of sleep on weekends for participants inline, which is `r mean(nlsy$sleep_wknd, na.rm = TRUE)`. <br>


```{r}
#Change the means to medians and re-render.

#| output: true
med_sleepwknd <- median(nlsy$sleep_wknd, na.rm = TRUE)
```

-   Participants' median number of hours of sleep on weekends is `r med_sleepwknd`. 
-   I am calculating the median number of hours of sleep on weekends for participants inline, which is `r median(nlsy$sleep_wknd, na.rm = TRUE)`. 


## Exercise 2.3 (Part 2)

```{r}
#Add a chunk with the table1 table from the slides. Practice extracting the output using inline_text() as in the slides.

#| label: tbl-descr
#| tbl-cap: "Descriptive statistics"
#| output-location: slide
table1 <- tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir)) |> 
  add_overall(last = TRUE)
table1

inline_text(table1, variable = "glasses", column = "stat_0",
            pattern = "{n}/{N} ({p}%)")

```

```{r}
#Edit or create a new table that has the median hours of sleep on weekends in it (in the overall sample).
#From that table, pull out the that value and include in the text using inline_text().

#| label: tbl-two
#| tbl-cap: "Descriptive statistics"
#| output-location: slide
table2 <- tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir, sleep_wknd))|> 
  add_overall(last = TRUE)
table2

med_hrs_sleep <- inline_text(table2, variable = "sleep_wknd", column = "stat_0") 

medsleepwknd_f <- inline_text(table2, variable = "sleep_wknd", 
                         column = "stat_2")
medsleepwknd_m <- inline_text(table2, variable = "sleep_wknd", 
                         column = "stat_1")
```

The overall median hours of sleep on weekends and IQR are (`r med_hrs_sleep`). <br>

Females and male participants have the same median hours of sleep on weekends (`r medsleepwknd_f`) vs. (`r medsleepwknd_m`).


